### 2013-10-08 reply-to-editors.txt
some modifs after review

Reviewer: 1
Comments to the Author
This manuscript describes a classification system for automatic creation MoA annotations and generation of drug repurposing hypotheses. The authors have developed an online database and released its source code to disseminate this data. The article is clearly laid out and well explained. The approach is novel and should be useful both to the bioinformatic community and a wider scientific audience, but several issues certainly require further clarification.

Major revisions
1) The authors describes FTC is built by integrating information coming from various sources and using ELK algorithm to classify the knowledge. However, I cannot find any clear and detail introduction in manuscript or supplement. This should be explained in detail, being one of the most distinctive features of the paper.

2) Result section 2, 2.1 is mostly methods benchmarking. I suggest those section should be renamed as “methodology” section. In addition, the detail classification description in supplementary (i.e. section 3, 3.1 and 3.2) should be wrote at manuscript. It would make it clearer how the methods are developed and tested.

3) In Figure 1, part C (Incorporation of databases), authors show ‘Prothrombin’ (P00734) is involved in ‘positive regulation of blood coagulation’ (GO_0030194). I tried to check this information from UniProt, but I cannot find any description there (http://www.uniprot.org/uniprot/P00734). How is this relationship module structured? This should be explained in detail.

4) How much of the different results between ATC and FTC? It should be listed and accessed via web.

Minor points:
1) Authors used ATC corpus to evaluate proposed FTC algorithm. The detailed evaluation results were shown on project’s web. I suggest the results should be listed at manuscript, not only at web.
2) 4 Gb of memory --> 4 GB of memory

Reviewer: 2
Comments to the Author
The authors present a taxonomy, the Functional Therapeutic Chemical Classification System, and a knowledge base to assist drug repurposing analyses. The mode of action of a drug, represented by Gene Ontology derived terms, forms the basis for the drug similarity. The authors evaluate their taxonomy by comparing it against the Anatomical Therapeutic Chemical Classification System, and show the usefulness of their system by applying it in a use case to find new drugs for Alzheimers disease.

The knowledge base and taxonomy is very well-documented, but requires deep knowledge about both the system and description logic in general in order to use. Also, presenting a single example use case with only literature evidence to back up the results is not very convincing for a research paper.

Major
Please show that the system can reproduce one known case of drug repurposing.

The authors should describe differences with their system to related efforts such as work by Hoehndorf, Dumontier, and Gkoutos. Bioinformatics 28.16 (2012): 2169-2175, and the Open PHACTS discovery platform http://www.openphacts.org/. A reference to a recent review paper such as the one by Andronis et al in Brief Bioinform. 2011 Jul;12(4):357-68 would be helpful as well.

In addition, the following specific revisions are needed:
p. 2, line 39: “In total there are 1280 approved DrugBank compounds (chemical and biotherapeutics) perturbing 1264 human protein targets.” I looked up the current Drugbank figures: the database contains 6811 drug entries including 1528 FDA-approved small molecule drugs, 150 FDA-approved biotech (protein/peptide) drugs, 87 nutraceuticals and 5080 experimental drugs. It is therefore not clear to me where the number 1280 comes from. Later the authors talk about “approved” drugs. Do they refer to FDA-approved?

p. 4 line 47: “Note that the mean of the similarity values was considered for the statistical analysis; some outliers are also present in the map, which can be interpreted as repurposing hypotheses.” Is this the only interpretation?

p. 4 line 51: “Five FTC categories containing drugs are directly related to the biological processes of the neurodegenerative condition:..” It seems like a user needs to know a lot of details about the system with > 23000 categories in order to use it. How can the user be helped here?

Minor

p. 2 line 26: “Drugs are further assigned into MoA categories based of the OWL constructs and axioms defined in the FTC (see Availability and Implementation).” Provide an example here.

p. 5 line 52: “The presence of lithium (group D) was unexpected and suspected of being a false positive at first glance.” Please explain why this result was unexpected.

Supplemental material:
“The web application built on the top of the FTC can be find …” should be “FTC can be found”

“The FTC implementation relies mostly on Brain [Croset et al., 2013] and the web application builds on the top of the Play! framework”. Please provide reference for the Play! Framework.