### 2013-10-10 reply-to-editors.txt
letter to editors done

Dear Editors,

We read with care the comments provided by the reviewers. We hereby present a modified version of the manuscript, based on their valuable input. Reviewer 1 asked for a more detailed presentation of the reasoning algorithm and methodology. In this regard, we have provided a supplementary reference to an introductory text. It is challenging to add more information to the main document, as reasoning services and tools have been developed by other research groups and are simply used in the context of our work. We rely on the W3C OWL international standard to interact with reasoning engines, for which extensive documentation is available. Moreover, we specially wrote an introductory text to description logics, available on the wiki for biomedical users non familiar with the domain. Ideally description logics would be presented in greater details in the main manuscript, unfortunately and because of space concerns we decided to focus on the presentation of the biomedical novelty brought by our project. Understanding in details the methodology requires some knowledge about description logics (as noticed by Reviewer 2 too), yet we believe that Figure 1 provides a clear simplified overview. Reviewer 1 did not manage to find the Gene Ontology (GO) annotation "positive regulation of blood coagulation" for the protein Prothrombin (P00734). The annotation exists and is present on the GO annotations website: http://www.ebi.ac.uk/QuickGO/GProtein?ac=P00734. Reviewer 1 understood correctly the methodology, yet probably missed the link "Complete GO annotation..." on the Uniprot website, leading to the complete list where the mentioned annotation can be found. Reviewer 1 also asks for "how much of the different results between ATC and FTC". The relation between the two resources is described in section 2.1 "Evaluation" and furthermore on the website of the application (https://www.ebi.ac.uk/chembl/ftc/evaluation/), where we compare equivalent classes and derive metrics (precision and recall) summarising the overlap between these resources. It is not clear to us what are the specific requirements of reviewer 1 regarding this revision.
Reviewer 2 asked us to illustrate the potential of the resource with a known repurposing example. We chose the thalidomide to fulfil this legitimate requirement, as it is possibly one of the best known repositioning case. We have added a paragraph to the manuscript (end of section 2.2.1) explaining how the teratogenic effect of the molecule is described in the FTC and how the new indications of the drug are captured (anti-cancer and immunomodulatory activities, nowadays routinely prescribed to patients). We briefly discussed the relationship between our resource and the work done in Bioinformatics 28.16 (2012): 2169-2175 and the Open PHACTS platform. Moreover, the methodology section further references an omitted similar approach presented in Nucleic Acids Res. 2011 Oct;39(18):e119. Reviewer 2 pointed the confusing number of drugs and their origin. This section was clarified in the manuscript (section 2); we used the FDA-approved drugs as described in DrugBank and interacting with at least one human protein target. Reviewer 2 asked whether the outliers points present in Figure 4 could have different interpretation; the text was extended accordingly, and the meaning of the data points summarily discussed. Reviewer 2 expressed concerns regarding the user interaction with the resource, as a large number of new classes are present. We argue that the web application greatly improves the accessibility of the FTC's content. Each FTC category is described in details on the web application, and is further linked to related terms for the user to browse. Moreover, a text search functionality is available (https://www.ebi.ac.uk/chembl/ftc/search/), in order to easily find categories or drugs of interest. More search functionality might be added in the future, depending on the adoption of the classification.
Minor corrections have been incorporated, as the exception of supplementary tables listing all the evaluation points. We believe that the website summarises better this aspect of the work. The text of the manuscript was also extensively edited and proof-read. Finally, Reviewer 3 did not manifest any major or minor objections.

We believe to have addressed or answered all the major comments raised by the referees with this new improved version of the manuscript. We are looking forward to hear your decision.

Best regards,

Samuel Croset